---
title: "R4DS Study Group - Week 1"
author: "Pierrette Lo"
date: "4/7/2020"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## This week's assignment

* Chapter 1 (Welcome - Introduction)
* Chapter 2 (Explore - Introduction)
* Chapter 3.1 - 3.2 (First Steps)

### Notes

On typing code:

* It's good practice to type in code yourself, rather than just copying & pasting
* Certain things will autofill, e.g. (), "", etc. while you're typing
* Tooltips will show when you start to type a function
* Pop-up suggestions may also show -- you can select using up/down arrow and Enter
* You can insert line breaks in your code to make it more readable. Hit Enter to start a new line, and the prompt will change from > to +

RStudio "dark mode":

* Tools > Global options > Appearance

R vocabulary:

* Dataframe = table; tibble = dataframe with special features
* Objects and variables = nouns
* Functions = verbs

How to read a Help page:

* e.g. `?mpg`
* Title: short description
* Description: more detail
* Usage: what to type
* Format (for datasets): what is in the dataset
* Arguments (required vs default): what are the options you can use with the function
* Details: more about how the function works
* Value: what is the output of the function
* Examples

### Following along with Chapter 3 text

Load tidyverse (you'll see several warning messages that can be ignored):
```{r, results="hide"}
library(tidyverse)
```

Take a look at the built-in `mpg` dataset:
```{r}
mpg
```

Look at the help page for `mpg`:
```{r, results="hide", message=F, eval=F}
?mpg
```

Run the code to make a plot:
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

### Chapter 3 exercises

See https://jrnold.github.io/r4ds-exercise-solutions/data-visualisation.html for more detailed solutions.

1. Run `ggplot(data = mpg)`. What do you see?
```{r}
ggplot(data = mpg)
```

Just an empty plot - the other components haven't been specified yet.

2. How many rows are in `mpg`? How many columns?
```{r}
mpg
```

Typing name of dataset - first line of output tells you number of rows x columns

Or you can try the below functions to summarize or count rows and columns

```{r, results="hide"}
structure(mpg)
glimpse(mpg)
nrow(mpg)
ncol(mpg)
```

3. What does the `drv` variable describe? Read the help for `?mpg` to find out.
```{r, results="hide", message=F, eval=F}
?mpg
```

drv = drive system of car

Categorical variable = finite number of categories/groups, not necessarily ordered

Continuous variable = infinite number of possible values (e.g. numbers, time)

4. Make a scatterplot of `hwy` vs `cyl`.
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = hwy, y = cyl))
```

5. What happens if you make a scatterplot of `class` vs `drv`? Why is the plot not useful?
```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = class, y = drv))
```

Both `class` and `drv` are categorical - the scatterplot only tells you when the categories overlap, but not how many datapoints are in each. 

See solutions manual for examples of better ways to visualize 2 categorical variables.
